### 2024-10-22 main_widom_morse.tex
add description of Fig2a, Fig2b in text

 In the supercritical region, the substance no longer behaves distinctly as either a liquid or a gas. Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary. Nevertheless, fluids exhibit distinct liquid-like and gas-like structures beyond the critical temperature. A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010}. This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances. Although, the equation of state is derived in the mean-field-like approximation,   in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}). It means the possibility to plot the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
 In what follows we consider this in more detail.

The isothermal compressibility $\kappa_T$ is defined by
\begin{equation}
	\label{def:isotherm_compres}
	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
\end{equation}
In \cite{DobushEtAl2024} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced} 
It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
\begin{eqnarray}
	\label{eq:kappa_star_m1}
	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\\
	\label{eq:kappa_star_m}
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
The reduced particle number density $\rho^*$ is found by 
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020,DobushEtAl2024}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
Finally, derive the isothermal compressibility in explicit form  
\begin{eqnarray}
	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
\end{eqnarray}

@@

[Figures only. Spots of text in progress.]

\begin{figure}[h!] 
	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The two figures differ in the scale of $M$. Part~(a) covers a wider range of $M$. Part~(b) focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
	}
	\label{fig3}
\end{figure}

>>>>

 In the supercritical region, the substance no longer behaves distinctly as either a liquid or a gas. Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary. Nevertheless, fluids exhibit distinct liquid-like and gas-like structures beyond the critical temperature. A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010}. This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances. Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,   in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}). It means the possibility to plot the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
 In what follows we consider this in more detail.

The isothermal compressibility $\kappa_T$ is defined by
\begin{equation}
	\label{def:isotherm_compres}
	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
\end{equation}
In \cite{DobushEtAl2024} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced} 
It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
\begin{eqnarray}
	\label{eq:kappa_star_m1}
	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\\
	\label{eq:kappa_star_m}
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
The reduced particle number density $\rho^*$ is found by 
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020,DobushEtAl2024}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
Finally, derive the isothermal compressibility in explicit form  
\begin{eqnarray}
	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
\end{eqnarray}
\begin{figure}[h!] 
	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The two figures differ in the scale of $M$. Part~(a) covers a wider range of $M$. Part~(b) focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
	}
	\label{fig3}
\end{figure}
Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3}a (wide range of $M$), and Figure~\ref{fig3}b (narrow range around $M=0$). Right at the critical point the compressibility diverge. As a result, there is still a "memory" of the critical point, where $\kappa_T$ reach local maxima, but once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases. 
Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities. Above the critical point, the fluid becomes continuous. However, Figure~\ref{fig2}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region. In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior. In the second case ($M>0$), it behaves like a liquid. Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$. At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig2}b), they completely vanish at $\tau > 0.28$.

@@

[Figure only. Spots of text in progress.]
